#Download data  
```{r, echo = TRUE, results = 'hide'}
# Reproducible Research Peer Assessment #1

# Create drectory to save data
if (!file.exists("./RepData/data")) dir.create("./RepData/data")

#Download zip file and unzip it if it is not in data folder
if (!file.exists("./RepData/data/repdata-data-activity.zip")) {
  
  fileURL <- "http://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
  download.file(url = fileURL, destfile = "./RepData/data/repdata-data-activity.zip")
  
  unzip("./RepData/data/repdata-data-activity.zip", exdir = "./RepData/data")
}
```
    
#Read in data  
```{r, echo = TRUE}
act <- read.csv('./RepData/data/activity.csv', stringsAsFactors = F)  #17568 obs, 3 cols
act$date <-as.Date(act$date, '%Y-%m-%d')
str(act)
head(act)
tail(act)
```
    
##Q1: What is mean total number of steps taken per day?  
```{r, echo = TRUE}
steps_per_day <- aggregate(x = act$steps, by = list(date = act$date), FUN = sum, na.rm = T)
hist(steps_per_day$x, xlab = 'Total number of steps per day', main = 'Total number of steps taken per day')
```


```{r, echo = TRUE}
mean = round(mean(steps_per_day$x),2)
median = median(steps_per_day$x)
```
The mean and median of total number of steps taken per day are `r mean` and `r median` respectively.  
    
##Q2: What is the average daily activity pattern?  
```{r, echo = TRUE}
steps_per_int <- aggregate(act$steps, by = list(Interval = act$interval), FUN = mean, na.rm = T)
Interval_Max = with(steps_per_int, Interval[x == max(x)])
plot(steps_per_int$Interval, steps_per_int$x, xlab = '5-minute interval', ylab = 'Average number of steps taken', type = 'l')
abline(v = Interval_Max, col ='blue')
```

The number of steps peaks at the `r Interval_Max`-min interval. People were more active around 8:30 am.  
    
##Q3: Imputing missing values 
```{r, echo = TRUE}
steps_miss <- sum(is.na(act$steps))
```
There are `r steps_miss` missing values in the dataset.   

    
The following code uses the mean of 5-min interval to impute number of steps missing in that interval:
```{r, echo = TRUE}
library(sqldf)
act_impute <- sqldf('select case when a.steps is null then b.x
                                 else a.steps end as steps_imp,
                    a.date, a.interval, b.x
                    from act a left join steps_per_int b on a.interval = b.Interval')
str(act_impute)

#Total number of steps taken each day
steps_per_day <- with(act_impute, aggregate(steps_imp, by = list(date = date), FUN = sum))

#Histogram
hist(steps_per_day$x, xlab = 'Total number of steps per day', main = 'Histogram of imputed total number of steps taken per day')

#Mean and median of total number of steps taken each day
options(scipen=999)
mean = round(mean(steps_per_day$x), digits = 0)
median = round(median(steps_per_day$x), digits =0)
```
The mean and median of total number of steps taken per day are both equal to `r mean`. Imputing the missing values with the mean per interval makes the mean and median closer to each other, i.e. distribution more symmetric.
    
    
##Q4: Are there differences in activity patterns between weekdays and weekends?

```{r, echo = TRUE}
#Weekend/weekday indicator
act$wkend <- ifelse(weekdays(act$date, abbreviate = T) %in% c('Sat', 'Sun'), 'Weekend', 'Weekday')
table(act$wkend, weekdays(act$date, abbreviate = T))

#Average # steps per interval over all weekday days and all weekend days
steps_per_int <- aggregate(steps ~ wkend + interval, data = act, FUN = "mean", na.rm = T)

#Stacking time series plot
library(ggplot2)
g <- ggplot(data = steps_per_int, aes(x = interval, y = steps))
g + geom_line(color = 'dark green', size = 1) +
  facet_grid(wkend ~.) +
  labs(x ='5-min interval', y = 'Number of steps', 
       title = 'Activity pattern during weekdays vs weekends')
```
